---
title: "Figure 2. Root microbiota structure shifts with increasing tiller number under field conditions"
author: "Yong-Xin Liu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo=T, comment=NA, message=F, warning=F,
	fig.align="center", fig.width=5, fig.height=3, dpi=300)
# Clean workspace
rm(list=ls()) 
source("http://210.75.224.110/stat_plot_functions.R")
w=89
h=59
```

## 2A. Density plot showing the distribution of the tiller number of 165 field-grown rice varieties

```{r LN_tiller}
phenoLN = read.table("../data/cultivar_LN_pheno.txt", header=T, row.names=1, sep="\t", comment.char="", stringsAsFactors=F)
phenoLN$soiltype = "LN"
pheno = phenoLN
step = 5
vline = sort(unique(round(pheno$tiller/step)*step+step*0.5 + 0.5))
vline = vline[1:length(vline)-1]

library(ggpubr)
(p = ggdensity(pheno, x="tiller", rug = TRUE, color = "soiltype", 
               fill = "soiltype",palette = c("#00AFBB", "#E7B800")) + 
    geom_vline(xintercept = vline,linetype=4)+ 
    xlab("Tiller number")+ylab("Percentage (%)") + 
    scale_x_continuous(breaks=vline, labels=vline) +
    main_theme +theme(legend.position="none")) # add = "mean", 
ggsave(paste("2a.tiller_LN_density.pdf", sep=""), p, width = w, height = h, units = "mm")
```


## 2B. Constrained PCoA with Bray–Curtis distance showing that rice varieties in each tiller number range harbored distinct root microbiota

```{r}
metadata = read.table("../data/cultivar_LN_metadata.txt", header=T, row.names=1, sep="\t", comment.char="", stringsAsFactors=F)
# set factor level
level=c("<8","[8,13)","[13,18)","[18,23)","[23,28)",">=28")
metadata$group = factor(metadata$group, level)

sub_design = metadata
m = "bray"
beta = read.table(paste("../data/cultivar_LN_otutab_norm.txt",sep=""), header=T, row.names=1, sep="\t", comment.char="") 

# Extract only those samples in common between the two tables
idx = rownames(sub_design) %in% colnames(beta)
sub_design=sub_design[idx,]
sub_beta=beta[,rownames(sub_design)]
otutab = t(sub_beta)/colSums(sub_beta,na=T)*100

# Constrained analysis OTU table by genotype
capscale.gen = capscale(otutab ~ group, data=sub_design, add=F, sqrt.dist=T, distance= m) 

# ANOVA-like permutation analysis
perm_anova.gen = anova.cca(capscale.gen, permutations = 1000, parallel = 9)

# generate variability tables and calculate confidence intervals for the variance
var_tbl.gen = variability_table(capscale.gen)
eig = capscale.gen$CCA$eig
variance = var_tbl.gen["constrained", "proportion"]
p.val = perm_anova.gen[1, 4]

# extract the weighted average (sample) scores
points = capscale.gen$CCA$wa[, 1:2]
points = as.data.frame(points)

points = cbind(points, sub_design$group)
colnames(points) = c("PC1", "PC2", "group")
points$PC1 = points$PC1 * -1
points$PC2 = points$PC2 * -1 

# plot PC 1 and 2
p = ggplot(points, aes(x=PC1, y=PC2, color=group)) + geom_point(alpha=.7, size=2) +
	labs(x=paste("CPCo1 (", format(100 * eig[1] / sum(eig), digits=4), "%)", sep=""),
	y=paste("CPCo2 (", format(100 * eig[2] / sum(eig), digits=4), "%)", sep="")) + 
	ggtitle(paste(format(100 * variance, digits=3), " % of variance; P = ",format(p.val, digits=2),sep="")) + 
	theme_classic() + main_theme
p = p + stat_ellipse(level=0.68,aes(fill = group), geom = "polygon", alpha = 0.2)
p

# save data table
write.table("Axis\t", file=paste("2b_cpcoa.txt",sep = ""), append = F, sep="\t", quote=F,  eol = "",row.names=F, col.names=F)
suppressWarnings(write.table(points, file=paste("2b_cpcoa.txt",sep = ""), append = T, sep="\t", quote=F, row.names=T, col.names=T))


arrows_pos = points %>% group_by(group) %>% summarise_all(mean)
arrows_pos = as.data.frame(arrows_pos)
n = dim(arrows_pos)[1]
arrows_pos = cbind(arrows_pos[1:(n-1),2:3], arrows_pos[2:n,2:3])
colnames(arrows_pos) = c("x","y","xend","yend")
p = p + geom_segment(data = arrows_pos, aes(x = x, y = y, xend = xend, yend = yend),
                  arrow = arrow(length = unit(0.2, "cm")),color="black",alpha=0.6)
p = p + theme(legend.position = "bottom")
p
ggsave(paste("2b.LN_beta_cpcoa_tiller.pdf", sep=""), p, width = w, height = h*1.3, units = "mm")
```

## 2C. OTU richness of root microbiota in rice plants within each tiller number range

```{r}
library(amplicon)
alpha = read.table(paste("../data/cultivar_LN_alpha.txt",sep=""), header=T, row.names=1, sep="\t", comment.char="") 
design = read.table("../data/cultivar_LN_metadata.txt", header=T, row.names=1, sep="\t")

# fiLter data
sub_design = design
idx = rownames(sub_design) %in% rownames(alpha)
sub_design=sub_design[idx,]
alpha=alpha[rownames(sub_design),]

method = c("richness","shannon_e") 
for(m in method){
  alpha_index = m # "richness"
  (p = alpha_boxplot(alpha, index = alpha_index, sub_design, groupID = "group") + theme(legend.position = "none"))
  ggsave(paste("2c.LN_alpha_", alpha_index, "_tiller_group6.pdf", sep=""), p, width = w, height = h, units = "mm")
}
```

## 2D. Phylum-level distribution of root microbiota in each tiller number range. Proteobacteria are shown at the class level

```{r}
metadata = read.table("../data/cultivar_LN_metadata.txt", header=T, row.names=1, sep="\t", comment.char="", stringsAsFactors=F)
level=c("<8","[8,13)","[13,18)","[18,23)","[23,28)",">=28")
metadata$group = factor(metadata$group, level)

tax_phylum = read.table(paste("../data/cultivar_LN_taxj_pc.txt",sep=""), header=T, row.names=1, sep="\t", comment.char="") 

library(amplicon)
(p = tax_stackplot(tax_phylum, metadata, topN = 8, groupID = "group", style = "group", sorted = "abundance"))
p = p + theme(legend.position = "bottom")
ggsave(paste("2d.LN_tax_PC_tiller.pdf", sep=""), p, width = w, height = h*1.3, units = "mm")


# tax_stackplot #----
tax_sum = tax_phylum
metadata = metadata
topN = 8
groupID = "group"
style = "group"
sorted = "abundance"

p_list = c("ggplot2", "reshape2")
for (p in p_list) {
    if (!requireNamespace(p)) {
        install.packages(p)
    }
    library(p, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
idx = rownames(metadata) %in% colnames(tax_sum)
metadata = metadata[idx, ]
tax_sum = tax_sum[, rownames(metadata)]
sampFile = as.data.frame(metadata[, groupID], row.names = row.names(metadata))
colnames(sampFile)[1] = "group"
mean_sort = as.data.frame(tax_sum[(order(-rowSums(tax_sum))), ])
other = colSums(mean_sort[topN:dim(mean_sort)[1], ])
mean_sort = mean_sort[1:(topN - 1), ]
mean_sort = rbind(mean_sort, other)
rownames(mean_sort)[topN] = c("Other")
merge_tax = mean_sort

# Correlation with sample RA with tiller ranges
group.cor = as.data.frame(cor(t(mean_sort), as.numeric(gsub("T","",metadata$tiller_cat)), method = "spearman"))
library(psych)
p = corr.test(t(mean_sort), as.numeric(gsub("T","",metadata$tiller_cat)), method = "spearman")  
group.cor2 = cbind(group.cor, p$p)
colnames(group.cor2)=c("Cor","Pvalue")
write.table("OTUID\t", file=paste("2d.LN_phylum_tax_cor_sample.txt",sep = ""), append = F, sep="\t", quote=F,  eol = "",row.names=F, col.names=F)
suppressWarnings(write.table(round(group.cor2,digits = 3), file=paste("2d.LN_phylum_tax_cor_sample.txt",sep = ""), append = T, sep="\t", quote=F, row.names=T, col.names=T))

mat_t = t(merge_tax)
mat_t2 = merge(sampFile, mat_t, by = "row.names")
mat_t2 = mat_t2[, c(-1)]
mat_mean = aggregate(mat_t2[, -1], by = mat_t2[1], FUN = mean)
mat_mean_final = do.call(rbind, mat_mean)[-1, ]
geno = mat_mean$group
colnames(mat_mean_final) = geno
mean_sort = as.data.frame(mat_mean_final)
mean_sort$Taxonomy = rownames(mean_sort)
data_all = as.data.frame(melt(mean_sort, id.vars = c("Taxonomy")))

# Correlation with group mean RA with tiller ranges
group.cor = as.data.frame(cor(t(mean_sort[,1:6]), 1:6, method = "spearman"))
library(psych)
p = corr.test(t(mean_sort[,1:6]), 1:6, method = "spearman")  #计算p值
# p$p
group.cor2 = cbind(group.cor, p$p)
colnames(group.cor2)=c("Cor","Pvalue")
write.table("OTUID\t", file=paste("2d.LN_phylum_tax_cor.txt",sep = ""), append = F, sep="\t", quote=F,  eol = "",row.names=F, col.names=F)
suppressWarnings(write.table(round(group.cor2,digits = 3), file=paste("2d.LN_phylum_tax_cor.txt",sep = ""), append = T, sep="\t", quote=F, row.names=T, col.names=T))
```

### S4A. Density plot showing the distribution of tiller number for 165 rice varieties grown in the other test field

```{r HN_tiller}
phenoHN = read.table("../data/cultivar_HN_pheno.txt", header=T, row.names=1, sep="\t", comment.char="", stringsAsFactors=F)
phenoHN$soiltype = "HN"
pheno = phenoHN
step = 7
vline = sort(unique(round((pheno$tiller-5)/step)*step+step*0.5+5+0.5))
vline = vline[1:length(vline)-1]
library(ggpubr)
(p = ggdensity(pheno, x="tiller", rug = TRUE, color = "soiltype", 
               fill = "soiltype",palette = c("#00AFBB", "#E7B800")) + 
    geom_vline(xintercept = vline,linetype=4)+ 
    xlab("Tiller number")+ylab("Percentage (%)") + 
    scale_x_continuous(breaks=vline, labels=vline) +
    main_theme +theme(legend.position="none")) # add = "mean", 
ggsave(paste("S2a.tiller_HN_density.pdf", sep=""), p, width = w, height = h, units = "mm")
```


### S4B. Constrained PCoA with Bray–Curtis distance showing that rice varieties in each tiller number range harbored distinct root microbiota

```{r}
metadata = read.table("../data/cultivar_HN_metadata.txt", header=T, row.names=1, sep="\t", comment.char="", stringsAsFactors=F)
level=c("<16","[16,23)","[23,30)","[30,37)","[38,45)",">=45")
metadata$group = factor(metadata$group, level)

sub_design = metadata
m = "bray"
beta = read.table(paste("../data/cultivar_HN_otutab_norm.txt",sep=""), header=T, row.names=1, sep="\t", comment.char="") 

# Extract only those samples in common between the two tables
idx = rownames(sub_design) %in% colnames(beta)
sub_design=sub_design[idx,]
sub_beta=beta[,rownames(sub_design)]
otutab = t(sub_beta)/colSums(sub_beta,na=T)*100

# Constrained analysis OTU table by genotype
capscale.gen = capscale(otutab ~ group, data=sub_design, add=F, sqrt.dist=T, distance= m) 

# ANOVA-like permutation analysis
perm_anova.gen = anova.cca(capscale.gen, permutations = 1000, parallel = 9)

# generate variability tables and calculate confidence intervals for the variance
var_tbl.gen = variability_table(capscale.gen)
eig = capscale.gen$CCA$eig
variance = var_tbl.gen["constrained", "proportion"]
p.val = perm_anova.gen[1, 4]

# extract the weighted average (sample) scores
points = capscale.gen$CCA$wa[, 1:2]
points = as.data.frame(points)

points = cbind(points, sub_design$group)
colnames(points) = c("PC1", "PC2", "group")
points$PC1 = points$PC1 * -1 

# plot PC 1 and 2
p = ggplot(points, aes(x=PC1, y=PC2, color=group)) + geom_point(alpha=.7, size=2) +
	labs(x=paste("CPCo1 (", format(100 * eig[1] / sum(eig), digits=4), "%)", sep=""),
	y=paste("CPCo2 (", format(100 * eig[2] / sum(eig), digits=4), "%)", sep="")) + 
	ggtitle(paste(format(100 * variance, digits=3), " % of variance; P = ",format(p.val, digits=2),sep="")) + 
	theme_classic() + main_theme
p = p + stat_ellipse(level=0.68,aes(fill = group), geom = "polygon", alpha = 0.2)
p

# save data table
write.table("Axis\t", file=paste("S4b_cpcoa.txt",sep = ""), append = F, sep="\t", quote=F,  eol = "",row.names=F, col.names=F)
suppressWarnings(write.table(points, file=paste("S4b_cpcoa.txt",sep = ""), append = T, sep="\t", quote=F, row.names=T, col.names=T))

arrows_pos = points %>% group_by(group) %>% summarise_all(mean)
arrows_pos = as.data.frame(arrows_pos)
n = dim(arrows_pos)[1]
arrows_pos = cbind(arrows_pos[1:(n-1),2:3], arrows_pos[2:n,2:3])
colnames(arrows_pos) = c("x","y","xend","yend")
p = p + geom_segment(data = arrows_pos, aes(x = x, y = y, xend = xend, yend = yend),
                  arrow = arrow(length = unit(0.2, "cm")),color="black",alpha=0.6)
p = p + theme(legend.position = "bottom")
p
ggsave(paste("S4b.HN_beta_cpcoa_tiller.pdf", sep=""), p, width = w, height = h*1.3, units = "mm")
```

### S4C. OTU richness of root microbiota in rice plants within each tiller number range

```{r}
library(amplicon)
alpha = read.table(paste("../data/cultivar_HN_alpha.txt",sep=""), header=T, row.names=1, sep="\t", comment.char="") 
design = read.table("../data/cultivar_HN_metadata.txt", header=T, row.names=1, sep="\t")

# fiLter data
sub_design = design
idx = rownames(sub_design) %in% rownames(alpha)
sub_design=sub_design[idx,]
alpha=alpha[rownames(sub_design),]

method = c("richness","shannon_e") 
for(m in method){
  alpha_index = m # "richness"
  (p = alpha_boxplot(alpha, index = alpha_index, sub_design, groupID = "group")+main_theme + theme(legend.position = "none")+ 
    xlab("Ranges of tiller number")
    )
  ggsave(paste("S4c.HN_alpha_", alpha_index, "_tiller_group6.pdf", sep=""), p, width = w, height = h, units = "mm")
}
```

### S4D. Phylum-level distribution of root microbiota in each tiller number range grown in other test field

```{r}
metadata = read.table("../data/cultivar_HN_metadata.txt", header=T, row.names=1, sep="\t", comment.char="", stringsAsFactors=F)
level=c("<16","[16,23)","[23,30)","[30,37)","[38,45)",">=45")
metadata$group = factor(metadata$group, level)

tax_phylum = read.table(paste("../data/cultivar_HN_taxj_pc.txt",sep=""), header=T, row.names=1, sep="\t", comment.char="") 

library(amplicon)
(p = tax_stackplot(tax_phylum, metadata, topN = 8, groupID = "group", style = "group", sorted = "abundance"))
p = p + xlab("Ranges of tiller number") + theme(legend.position = "bottom")
ggsave(paste("S4d.HN_tax_PC_tiller.pdf", sep=""), p, width = w, height = h*1.3, units = "mm")

# tax_stackplot #----
tax_sum = tax_phylum
metadat = metadata
topN = 8
groupID = "group"
style = "group"
sorted = "abundance"

p_list = c("ggplot2", "reshape2")
for (p in p_list) {
    if (!requireNamespace(p)) {
        install.packages(p)
    }
    library(p, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}
idx = rownames(metadata) %in% colnames(tax_sum)
metadata = metadata[idx, ]
tax_sum = tax_sum[, rownames(metadata)]
sampFile = as.data.frame(metadata[, groupID], row.names = row.names(metadata))
colnames(sampFile)[1] = "group"
mean_sort = as.data.frame(tax_sum[(order(-rowSums(tax_sum))), 
    ])
other = colSums(mean_sort[topN:dim(mean_sort)[1], ])
mean_sort = mean_sort[1:(topN - 1), ]
mean_sort = rbind(mean_sort, other)
rownames(mean_sort)[topN] = c("Other")
merge_tax = mean_sort

group.cor = as.data.frame(cor(t(mean_sort), as.numeric(gsub("T","",metadata$tiller_cat)), method = "spearman"))
library(psych)
p = corr.test(t(mean_sort), as.numeric(gsub("T","",metadata$tiller_cat)), method = "spearman")  
group.cor2 = cbind(group.cor, p$p)
colnames(group.cor2)=c("Cor","Pvalue")
write.table("OTUID\t", file=paste("S4d.HN_phylum_tax_cor_sample.txt",sep = ""), append = F, sep="\t", quote=F,  eol = "",row.names=F, col.names=F)
suppressWarnings(write.table(round(group.cor2,digits = 3), file=paste("S4d.HN_phylum_tax_cor_sample.txt",sep = ""), append = T, sep="\t", quote=F, row.names=T, col.names=T))

mat_t = t(merge_tax)
mat_t2 = merge(sampFile, mat_t, by = "row.names")
mat_t2 = mat_t2[, c(-1)]
mat_mean = aggregate(mat_t2[, -1], by = mat_t2[1], FUN = mean)
mat_mean_final = do.call(rbind, mat_mean)[-1, ]
geno = mat_mean$group
colnames(mat_mean_final) = geno
mean_sort = as.data.frame(mat_mean_final)
mean_sort$Taxonomy = rownames(mean_sort)
data_all = as.data.frame(melt(mean_sort, id.vars = c("Taxonomy")))

group.cor = as.data.frame(cor(t(mean_sort[,1:6]), 1:6, method = "spearman"))
library(psych)
p = corr.test(t(mean_sort[,1:6]), 1:6, method = "spearman")  #计算p值
# p$p
group.cor2 = cbind(group.cor, p$p)
colnames(group.cor2)=c("Cor","Pvalue")
write.table("OTUID\t", file=paste("S4d.HN_phylum_tax_cor.txt",sep = ""), append = F, sep="\t", quote=F,  eol = "",row.names=F, col.names=F)
suppressWarnings(write.table(round(group.cor2,digits = 3), file=paste("S4d.HN_phylum_tax_cor.txt",sep = ""), append = T, sep="\t", quote=F, row.names=T, col.names=T))
```
